Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: updated detailed summary table styles for clarity

\begin{table}[t]
\begin{tabular}{c | c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
\hline
EdGB        & 4 & 0.0107 & 0.0044 & $-$0.2480 & $-$1.101 &  \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
dCS         & 4 & 3.1964 & 6.3619 & 41.199 & 794.66 &  \cite{Wagle:2021tam}  \\
cubic EFT   & 4 & -0.5813  & 2.6469 & -3.8620 & 265.12 & \cite{Cano:2021myl} \\
quartic EFT & 6 & \dots  & \dots & \dots & \dots & \cite{Cano:2021myl}  \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(i)}_{220}$ and decay time $\delta \tau^{(i)}_{220}$, the
references from which we used the results from and the current best constraint
(if applicable).
}
\label{tab:ref_theories_qnms}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    & No  & Yes & -- \\
GW150914 & $j = 0, 1$ & No  & Yes & -- \\
GW200129 & $j = 0$    & No  & Yes & -- \\
GW200129 & $j = 0, 1$ & No  & Yes & -- \\
Combined & $j = 0$    & --  & Yes & -- \\
Combined & $j = 0, 1$ & --  & Yes & -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for EdGB gravity for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We conclude that we cannot place any constraint on $\ell_{\rm GB}$ with any of the GW events
studied here.~\hs{TODO: update when completed.}
}
\label{tab:summary_edgb}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    & Yes & Yes & $\ell_{\rm CS} \leqslant 46.5$~km \\
GW150914 & $j = 0, 1$ & Yes & No  & --                                 \\
GW200129 & $j = 0$    & Yes & Yes & $\ell_{\rm CS} \leqslant 43.9$~km \\
GW200129 & $j = 0, 1$ & Yes & No  & --                                 \\
Combined & $j = 0$    & Yes & Yes & \cellcolor{blue!10}$\ell_{\rm CS} \leqslant 34.5$~km \\
Combined & $j = 0, 1$ & Yes & No  & --                                 \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for dCS gravity for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We found that while our posteriors satisfy the condition~\eqref{eq:eft_bound} (with $\varepsilon = 1/2$),
they do not obey the condition~\eqref{eq:parspec_bound} for $j = 1$. This means
that our results for $j=0$ are the only ones we can confidently quote.
%
The combined bound, which is also quoted in Table~\ref{tab:ref_theories_qnms},
is $\ell_{\rm CS} \leqslant 34.5$~km.
}
\label{tab:summary_dcs}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    &     &     &    \\
GW150914 & $j = 0, 1$ &     &     &    \\
GW200129 & $j = 0$    &     &     &    \\
GW200129 & $j = 0, 1$ &     &     &    \\
Combined & $j = 0$    &     &     &    \\
Combined & $j = 0, 1$ &     &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the cubic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_ceft}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    &     &     &    \\
GW150914 & $j = 0, 1$ &     &     &    \\
GW200129 & $j = 0$    &     &     &    \\
GW200129 & $j = 0, 1$ &     &     &    \\
Combined & $j = 0$    &     &     &    \\
Combined & $j = 0, 1$ &     &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_qeft}
\end{table}

@@

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso},
whcih yields,
%
\begin{equation}
p_{\rm CS} = 4\,, \quad \delta \omega^{(0)}_{\rm CS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm CS} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be carried for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm CS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm CS} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB{} waveform model.
%
We remark that the alarmingly large values of $\delta \omega_{\rm CS}^{(1)}$ and $\delta \tau_{\rm CS}^{(1)}$
are compensated by the assumptions that $(\ell_{\rm CS} / M_f)^2$ and $\chi$ are much less than unity. These
are the assumptions use to calculate the QNM frequencies both in Ref.~\cite{Wagle:2021tam} and also in
other gravity theories considered here~\cite{Pierini:2021jxd,Cano:2021myl}.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.

>>>>

\begin{table}[t]
\begin{tabular}{c | c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
\hline
EdGB        & 4 & 0.0107 & 0.0044 & $-$0.2480 & $-$1.101 &  \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
dCS         & 4 & 3.1964 & 6.3619 & 41.199 & 794.66 &  \cite{Wagle:2021tam}  \\
cubic EFT   & 4 & -0.5813  & 2.6469 & -3.8620 & 265.12 & \cite{Cano:2021myl} \\
quartic EFT & 6 & \dots  & \dots & \dots & \dots & \cite{Cano:2021myl}  \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
}
\label{tab:ref_theories_qnms}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event &  EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       &  bound? & bound?  &                            \\
\hline
  & GW150914 & No  & Yes & -- \\
0 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
  & GW150914 & No  & Yes & -- \\
1 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for EdGB gravity for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We conclude that we cannot place any constraint on $\ell_{\rm GB}$ with any of the GW events
studied here.~\hs{TODO: update when completed.}
}
\label{tab:summary_edgb}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
      & GW150914 & Yes & Yes & $\ell_{\rm CS} \leqslant 46.5$~km \\
0     & GW200129 & Yes & Yes & $\ell_{\rm CS} \leqslant 43.9$~km \\
      & Combined & Yes & Yes & \cellcolor{blue!10}$\ell_{\rm CS} \leqslant 34.5$~km \\
\hline
      & GW150914 & Yes & No  & --                                 \\
1     & GW200129 & Yes & No  & --                                 \\
      & Combined & Yes & No  & --                                 \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for dCS gravity for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We found that while our posteriors satisfy the condition~\eqref{eq:eft_bound} (with $\varepsilon = 1/2$),
they do not obey the condition~\eqref{eq:parspec_bound} for $j = 1$. This means
that our results for $j=0$ are the only ones we can confidently quote.
%
The combined bound, which is also quoted in Table~\ref{tab:ref_theories_qnms},
is $\ell_{\rm CS} \leqslant 34.5$~km.
}
\label{tab:summary_dcs}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 &  &     &    \\
0 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
  & GW150914 &  &     &    \\
1 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the cubic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_ceft}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 &  &     &    \\
0 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
  & GW150914 &  &     &    \\
1 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_qeft}
\end{table}

@@

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso},
which yields,
%
\begin{equation}
p_{\rm CS} = 4\,, \quad \delta \omega^{(0)}_{\rm CS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm CS} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
We carry the same steps for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm CS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm CS} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB{} waveform model.
%
We remark that the alarmingly large values of $\delta \omega_{\rm CS}^{(1)}$ and $\delta \tau_{\rm CS}^{(1)}$
are compensated by the assumptions that $(\ell_{\rm CS} / M_f)^2$ and $\chi$ are much less than unity. These
are the assumptions use to calculate the QNM frequencies both in Ref.~\cite{Wagle:2021tam} and also in
other gravity theories considered here~\cite{Pierini:2021jxd,Cano:2021myl}.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.
